A Theory for Distributed Database Transactions

### 2019-11-04 2018_Transactions/articleTransactions.tex
Improve language and arguments

	This makes transactions suitable for more databases than just relational ones.
	The soundness of the theory has been established by means of formal proof.
	To show that the theory works, the theory is demonstrated by a prototype developed in Ampersand\footnote{\tt https://github.com/AmpersandTarski/Ampersand}.
	To demonstrate that the theory is about real transactions, this contribution defines semantics of conventional database transactions.
	To fulfill our promise, we also gives semantics of transactions on the semantic web.
\end{abstract}

@@

\section{Introduction}
\label{sct:Introduction}
	A distributed information system (DIS) serves its users with things such as
	registering data, guarding consistency, protecting that data,
	and providing it accurately and meaningfully for its intended purpose to various users.
	These users want predictable behaviour, even when they change a shared state,
	e.g. when reserving a seat in a global flight reservation system like Galileo or Amadeus.
	Although we cannot always avoid sharing a state,
	we usually can avoid sharing an entire database.
	Globally distributed applications such as Netflix and LinkedIn cannot afford to rely on a shared database for obvious reasons of performance and reliability.
	Many cloud applications use distributed storage models instead.
	Examples are event stores (e.g.\ Kafka), triple stores (e.g.\ semantic web), graph databases (e.g. Neo4J) and others.
	Many so-called NoSQL databases lack functionality for transactions,
	leaving the problems for the user to solve.
	As a result, architectural patterns exist for the purpose of doing transactions in a distributed setting.
	Some of these patterns simply circumvent the problem.
	Others will isolate transactionality into a smaller sub-problem.
	Still others introduce locking mechanisms to do the job.

	Our work is inspired by the observation that the merge of Git corresponds to the commit that we know from transactional databases.
	We show this correspondence by expressing the characteristic properties of transactions in a theory of time.
	For this purpose we build on a well-known theory of time~\cite{DBLP:journals/ipl/FosterCWZ18},
	which is generally applicable,
	well-established, and founded in a solid mathematical structure~\cite{Foster14}.
	This means we have to redefine Atomicity, Consistency, Isolation, and Durability in terms of events.
	Informally, {\em Atomicity} means that a sequence of primitive actions can be executed as though it were a single action.
	{\em Consistency} means that integrity rules are warranted.
	{\em Isolation} means that actions can be done without any knowledge of what other users are doing.
	{\em Durability} means that history does not change.

@@

\section{State}
\label{sct:State}
	Typically, distributed information systems are stateful.
	Every event in the history of a DIS may cause a different (local) state.
	(The distributed nature of a DIS means that states are local.)
	Let $\mathbb S$ be the set of all conceivable states.
	So we need a function $\id{state}:\mathbb E\rightarrow\mathbb S$ that assigns one state to every event $e$ in history $\id{prec}$.
	State $\id{state}(e)$ represents the state immediately after $e$ has happened.
	If the history $\id{prec}$ is obvious from the context, we may write $\id{state}(e)$ rather than $\id{state}(e)$.
	We assume an empty state, $\emptyset$, which is the state immediately before an initial event has happened.
	The global state of the system, if relevant, can be computed by $\{\id{state}(e) | e\in\id{current}(\id{prec})\}$.
	Locally we work with one specific current event, $e$, with $\id{state}(e)$ representing the local current state.

@@

	In his paper "Splitting Atoms Safely"~\cite{JONES2007109} Cliff Jones\footnote{%
	Jones is author of the Atomicity Manifesto and author of the formal method VDM.}
	describes the intended meaning of atomicity in natural language:
	``If an action is said to be executed “atomically”, it is assumed that it will not be affected by interference
	and that the environment will not be able to observe intermediate steps of the action in question.''
	Jones argues that interference must be tolerated, however.
	With shared state programs, he says, an action must achieve some required result even though its state can be changed by other interfering processes.

>>>>

	This makes transactions suitable for more databases than just relational ones.
	The soundness of the theory has been established by means of formal proof.
	To show that the theory works, the theory is demonstrated by a prototype developed in Ampersand\footnote{\tt https://github.com/AmpersandTarski/Ampersand}.
	To demonstrate that the theory is about real transactions, this contribution defines semantics of conventional database transactions.
\end{abstract}

@@

\section{Introduction}
\label{sct:Introduction}
	A distributed information system (DIS) serves its users with things such as
	registering data, guarding consistency, protecting that data,
	and providing it accurately and meaningfully for its intended purpose to various users.
	These users want predictable behaviour, even when they change a shared state,
	e.g. when reserving a seat in a global flight reservation system like Galileo or Amadeus.
	Although we cannot always avoid sharing a state,
	we usually can avoid sharing an entire database.
	Globally distributed applications such as Netflix and LinkedIn cannot afford to rely on a shared database for obvious reasons of performance and reliability.
	Many cloud applications use distributed storage models instead.
	Examples are event stores (e.g.\ Kafka), triple stores (e.g.\ semantic web), graph databases (e.g. Neo4J) and others.
	Many so-called NoSQL databases lack functionality for transactions,
	leaving the problems for the user to solve.
	As a result, many architectural patterns have emerged for the purpose of doing transactions in a distributed setting.
	Some of these patterns simply advise how to avoid or circumvent the problem.
	Others will isolate transactionality into a smaller sub-problem,
        downsizing the problem and hoping to deal with the sub-problem in a conventional transactional database.
	Still others introduce locking mechanisms to do the job.
        However, sometimes we must avoid locking for performance reasons and avoid a globally shared database for reasons of size.
        As is the case with large, globally distributed information systems such as Facebook, LinkedIn, and Netflix.
        In case we want full transactional behaviour in such systems, we have a problem.

        This problem seems to be solved by Git.
	Our work is inspired by the observation that the merge of Git corresponds to a transactional commit.
	We show this correspondence by expressing the characteristic properties of transactions in a theory of time.
	For this purpose we build on a well-known theory of time~\cite{DBLP:journals/ipl/FosterCWZ18},
	which is generally applicable,
	well-established, and founded in a solid mathematical structure~\cite{Foster14}.
	For expressing transactions in terms of events we have to redefine Atomicity, Consistency, Isolation, and Durability.
	Informally, {\em Atomicity} means that a sequence of primitive actions can be executed as though it were a single action.
	{\em Consistency} means that integrity rules are warranted.
	{\em Isolation} means that actions can be done without any knowledge of what other users are doing.
	{\em Durability} means that history does not change.

We define the history of a distributed information system as a directed acyclig graph $\la\mathcal E,\mathcal C\ra$ with
\begin{definition}
\begin{eqnarray}\mathcal C\subseteq\mathcal E\times\mathcal E\\
\mathcal C^+\cap I_{\mathcal E}=\emptyset
\end{eqnarray}
\end{definition}
A transaction is a set of events.
\begin{enumerate}
\item   Atomic\\
A transaction is atomic if it can be replaced by a single event without consequences for other parts of the system.
\item   Consistent\\
If the initial states of a transaction satisfy a given set of rules, we say that a transaction maintains consistency if its final state satisfies that same set of rules.
\item   Isolated\\
A transaction is isolated if its events are not shared by other transactions.
\item   Durable\\
A transaction is durable if it is part of every future history.
\end{enumerate}

@@

\section{State}
\label{sct:State}
	Typically, distributed information systems may have states.
%        Consider for example Twitter: the entire timeline on your smartphone represents the state.
	Every new event in the history of a DIS may cause a different state.
	The distributed nature of a DIS means that states are local.
	Let $\mathbb S$ be the set of all conceivable states.
	We need a function $\id{state}:\mathbb E\rightarrow\mathbb S$ that assigns one state to every event $e$ in history $\id{prec}$.
	State $\id{state}(e)$ represents the state immediately after $e$ has happened.
	If the history $\id{prec}$ is obvious from the context, we may write $\id{state}(e)$ rather than $\id{state}(e)$.
	We assume an empty state, $\emptyset$, which is the state immediately before an initial event has happened.
	The global state of the system, if deemed relevant, can be computed by $\{\id{state}(e) | e\in\id{current}(\id{prec})\}$.
	Locally we work with one specific current event, $e$, with $\id{state}(e)$ representing the local current state.

@@

	In his paper ``Splitting Atoms Safely''~\cite{JONES2007109} Cliff Jones\footnote{%
	Jones is author of the Atomicity Manifesto and author of the formal method VDM.}
	describes the intended meaning of atomicity in natural language:
	``If an action is said to be executed “atomically”, it is assumed that it will not be affected by interference
	and that the environment will not be able to observe intermediate steps of the action in question.''
	Jones argues that interference must be tolerated, however.
	With shared state programs, he says, an action must achieve some required result even though its state can be changed by other interfering processes.
